Parsing Postfix log files

### 2009-03-27 doc/logparser-implementation.tex
Merged check-regexs into logs2regexs, and updated the documentation.

A program is provided with \parsername{} to ease the process of creating
new rules from unrecognised log lines, based on the algorithm developed by
Risto Vaarandi for his \acronym{SLCT}~\cite{slct-paper}.  The differences
between the two algorithms will be outlined as part of the general
explanation below.  The core of the \acronym{SLCT} algorithm is quite
simple: programs generally create log lines by substituting variable words
into a fixed pattern, and analysis of the frequency with which each word
occurs can be used to determine whether the word is variable or part of the
fixed pattern.  This classification can be used to group similar log lines
and generate a regex to match each group of log lines.  There are 4 steps
in the algorithm:

@@

    \item [Build regexes.]  The words are reassembled to produce a regex
        matching the log line, and a counter is maintained for each regex.
        Contiguous sequences of \texttt{.+} in the newly assembled regexes
        are collapsed to a single \texttt{.+}; any resulting duplicate
        regexes are combined, and their counters added together.  If a
        regex's counter is lower than the threshold supplied by the user
        the regex is discarded; this second threshold is independent of the
        threshold used to differentiate between fixed and variable words,
        but once again 1\% of log lines is a good starting point.  The new
        regexes are displayed for the user to add to the database, either
        as new rules or merged into the regexes of existing rules; the
        counter for each regex is also displayed, giving the user an
        indication of how many of the log lines that regex should match.
        Discarding regexes will result in some of the log lines not being
        matched; when the rules have been augmented with the new regexes,
        the original log files should be parsed again, and any remaining
        unparsed log lines used as input to this utility.

@@

XXX MERGE logs2regexes WITH check-regexes.  CHECK IF SLCT DOES THIS STEP\@.

A second utility is also provided that reads a list of new regexes and the
input given to the first utility.  It tries to match each input log line
with each regex, counting the number of log lines that match each regex,
warning the user if an input log line is matched by more than one regex,
and additionally warning if an input log line is not matched by any regex.
It displays a summary of how many input log lines each regex matched,
comparing it to the expected number of matches; this provides the user with
an easy method of checking if the regexes produced by the first utility are
correctly matching the input log lines they are based upon.

These utilities are not intended to create perfect regexes, but they
greatly reduce the effort required to parse new or different log lines.

>>>>

A program is provided with \parsername{} to ease the process of creating
new rules from unrecognised log lines, based on the algorithm developed by
Risto Vaarandi for his \acronym{SLCT}~\cite{slct-paper}.  The differences
between the two algorithms will be outlined as part of the general
explanation below.  The core of the \acronym{SLCT} algorithm is quite
simple: programs generally create log lines by substituting variable words
into a fixed pattern, and analysis of the frequency with which each word
occurs can be used to determine whether the word is variable or part of the
fixed pattern.  This classification can be used to group similar log lines
and generate a regex to match each group of log lines.  There are five steps
in the algorithm:

@@

    \item [Build regexes.]  The words are reassembled to produce a regex
        matching the log line, and a counter is maintained for each regex.
        Contiguous sequences of \texttt{.+} in the newly assembled regexes
        are collapsed to a single \texttt{.+}; any resulting duplicate
        regexes are combined, and their counters added together.  If a
        regex's counter is lower than the threshold supplied by the user
        the regex is discarded; this second threshold is independent of the
        threshold used to differentiate between fixed and variable words,
        but once again 1\% of log lines is a good starting point.

    \item [Test the new regexes.]  The final step replaces keywords in the
        new regexes in the same way as \parsername{} does, and compiles
        each regex to check they are valid.  A match will be attempted
        between all of the new regexes and each of the unparsed log lines
        the regexes were built from; the user will be warned if a log line
        is not matched by any regex, or if a log line is matched by more
        than one regex.  The number of log lines each regex matches is
        counted, as is the number of log lines matched by all regexes,
        though a log line is counted once only, even if matched by more
        than one regex.

        This step is not performed by the original algorithm.

@@

The new regexes are displayed for the user to add to the database, either
as new rules or merged into the regexes of existing rules; also displayed
are the number of log lines each regex was expected to match, and the
number it actually matched, to help the user notice problems in the new
regexes.  Discarding regexes will result in some of the log lines not being
matched; when the rules have been augmented with the new regexes, the
original log files should be parsed again, and any remaining unparsed log
lines used as input to this utility.

This utility is not intended to create perfect regexes, but it greatly
reduces the effort required to deal with unrecognised log lines.
